Comparing variations in sediment transport capacity between subglacial and subaerial systems

### 2023-03-10 theory.tex
small bits.. and new fig

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, creates an imperative to examine the processes controlling sediment discharge from glacierized catchments.

@@

  In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:tau_t}
      \tau_t \propto w_c\, v^2,
    \end{equation}
  \end{linenomath*}
  % 
  where $w_c$ is the width of the channel.
  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:v}
      v = \frac{Q_w}{S},
    \end{equation}
  \end{linenomath*}
  % 
  where $Q_w$ is water discharge,  and the denominator is the channel's cross-sectional area $S$.

@@

  Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
  In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
  The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
  As a result, the subglacial channel size can evolve on a longer timescale than variations in water discharge \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon_text.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$. Text below panels shows different variables response to high and low water discharge ($Q_{wh}$ and $Q_{wl}$, respectively).
        Relationships between the variables presented here are given in Equations~\ref{eq:tau_t} and~\ref{eq:v}.
        Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

@@

  \section{Conclusions}

  The pressurized nature of subglacial channels means that changing water discharge is accommodated by altering water velocity and shear stress, upon which sediment transport depends.
  This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
  Conversely, in subaerial systems changing water discharge modulates the water velocity as well as the channel width over which sediment can be transported.

>>>>

  Glaciers expel large quantities of sediment from their termini \citep{hallet1996}.
  Changes in glacier dynamics, geomorphology, and hydrology  have prompted numerous  recent studies of  sediment transport processes in cold regions \citep[e.g.][]{zhang2022}.
  Recent papers examining increased sediment discharge form glacierized regions at both the glacier and regional scale \citep{costa2017,delaney2020,li2022,vergara2022}, create an imperative to examine the processes controlling sediment discharge from glacierized catchments. 
\ian{more details about paper... why have they not examined the processes more?}

@@

  In a transport-limited regime, sediment discharge responds to the sediment transport capacity or the amount of sediment that could be carried by the water. Sediment transport capacity depends on the shear stress between water and sediment it flows  over \citep{shields1936,meyer1948,engelund1967} and the width of the channel $w_c$ over which to mobilize sediment. The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through at channel so that 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:tau_t}
      \tau_t \propto w_c\, v^2,
    \end{equation}
  \end{linenomath*}
  % 
  where $w_c$ is the width of the channel.
  Following the conservation of mass, the velocity of the water flowing through a rectangular channel is given as 
  \begin{linenomath*}
    \begin{equation}
      \label{eq:v}
      v = \frac{Q_w}{S},
    \end{equation}
  \end{linenomath*}
  % 
  where $Q_w$ is water discharge,  and $S$ is the channel's cross-sectional area.

@@

  Yet, the response of water velocity to changing water discharge in subglacial channels is different than subaerial ones.
  In subglacial systems, the water is pressurized by the ice above \citep{shreve1972}.
  The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \citep{rothlisberger1972}.
  As a result, the subglacial channel size evolves over weeks or months, whereas water discharge varies over hours \citep[e.g.][]{iken1986,andrews2014,nanni2020}.
  In turn, changes in water discharge $Q_w$ are mainly accommodated with water velocity $v$, because the size of the channel $S$ is fixed (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
  It follows that greater water discharge $Q_w$ will cause the shear stress across the channel $\tau_t$ to respond to the water velocity $v$, as opposed to the subaerial system where width evolves as well.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Water velocity magnitude in the subglacial, $v_g$ , and subaerial ,$v_f$ , systems are shown by arrow length marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in subglacial and subaerial channels. The subglacial channel widths $w_{cg}$ remain steady, while the subaerial channel widths evolve $w_{cf}$.
        Note that in the subaerial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

@@

  \section{Conclusions}
  Contrary to subaerial channels that can alter their channel width and water velocity in response to changing water discharge,
   pressurized subglacial channels accommodate changing water discharge by altering water velocity and shear stress, upon which sediment transport depends.
  This occurs because the size of subglacial channels evolves slowly compared to variations in water discharge.
